---
title: "HW_2_Data_607"
author: "Natalie Kalukeerthie"
date: "2024-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – R and SQL

Part 1: Build Table

• Choose six recent popular movies.
• Ask at least five people that you know (friends, family, classmates, imaginary friends if
necessary) to rate each of these movies that they have seen on a scale of 1 to 5.

Part 2: Store data in SQL database

• Take the results (observations) and store them in the class MySQL database:

- Server name: cunydata607sql.mysql.database.azure.com
- Username / password: will be given to you in an email

Note: it is good practice to change your password. To do so, use this SQL command:
SET PASSWORD = ‘<your new password here>';

Part 3: Transfer data from SQL database to R dataframe

• Load the information from the SQL database into an R dataframe.
Part 4: Missing data strategy

• Implement an approach to missing data

• Explain why you decided to take the chosen approach

Note: consider that later in the course you will revisit this information you have collected and
will use it to implement a Recommender.

### Build Table and Store Data in SQL Database

A table has been created and stored in the cunydata607sql.mysql.database.azure.com called "popular_movie_poll".

### Transfer data from SQL database to R dataframe
```{r dataframe}

#load the data
library(tidyverse)
library(openintro)

movie_poll <- read.csv('https://raw.githubusercontent.com/nk014914/Data-607/main/popular_movie_poll.csv')

```

### Missing Data Strategy

For missing data in this exercise, I used the "0" as the NULL when a rater did not watch the movie, this is because ratings were on a scale of 1 to 5, so zero was not in use, and because in SQL, I could not use "N/A" as the datatype of the ratings was INT. I would rather not use a blank as there is a potential to cause an issue if using the data for calculations. 
