---
title: "HW_3_Data_607"
author: "Natalie Kalukeerthie"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – Character Manipulation and Data Processing


Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to the problems below.  You may work in a small group, but please submit separately with names of all group participants in your submission.

### 1) Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
#load the data
library(tidyverse)
library(openintro)
library(RMySQL)

college_majors <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')

#view the data
head(college_majors)

#use str_detect to see if pattern is in the data
str_detect(college_majors, '(DATA|STATISTICS)')

#find the pattern using grep
grep('DATA|STATISTICS',college_majors$Major, value = TRUE,ignore.case = TRUE)

```
We see that there are only 3 majors in the list that contains either "Data" or "statistics"

### 2) Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

From my understanding, I believe I am suppose to take the above string and print it as what it would look like as a vector?
```{r}
fruits <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'


list_fruits <- str_extract_all(string = fruits, pattern = '\\".*?\\"')

#removing [x] characters
items <- str_c(list_fruits[[1]], collapse = ', ')
str_glue('c({items})', items = items)
```

The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:
### 3) Describe, in words, what these expressions will match:

    (.)\1\1 - Matches the same character appearing 3 times in a row ex. aaa
    "(.)(.)\\2\\1" Matches a pair of characters to the same pair but backwards ex. appa
    (..)\1 Matches any two characters that repeat ex. abab
    "(.).\\1.\\1" A character, then any character, then the original character again, then any character, then the original character ex. qrqsq
    "(.)(.)(.).*\\3\\2\\1" three characters, followed by 0 or more characters of any kind, followed by the original 3 characters backwards.  ex. abcdrlmnopcba

### 4) Construct regular expressions to match words that:

```{r}

random_words <- c("apple", "keys", "america","high","tonight","onomonopia","window","door","cocoa","bucket","eye","leg","arm")

#Start and end with the same character.
str_subset(random_words, "^(.)((.*\\1$)|\\1?$)")

#Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
str_subset(random_words,"([A-Za-z][A-Za-z]).*\\1")

#Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
str_subset(random_words, "([a-z]).*\\1.*\\1")

```