---
title: "HW_5_Data_607"
author: "Natalie Kalukeerthie"
date: "2024-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – Tidying and Transforming Data

The chart above describes arrival delays for two airlines across five destinations. Your task is to:

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.
You’re encouraged to use a “wide” structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy
and transform your data.

(3) Perform analysis to compare the arrival delays for the two airlines.

(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative
descriptions of your data cleanup work, analysis, and conclusions. Please include in your
homework submission:
 The URL to the .Rmd file in your GitHub repository. and
 The URL for your rpubs.com web page.


```{r}
#load the data
library(tidyverse)
library(openintro)
library(dplyr)

airline_delays <- read.csv('https://raw.githubusercontent.com/nk014914/Data-607/main/airline_destinations_data_untidy.csv')

#view the data
head(airline_delays)

```

```{r}
#Removing blank row and renaming columns
airline_delays <- airline_delays %>% 
  drop_na() %>%
  rename(airline = X, status = X.1)

#Filling NAs for missing airline cells then filling with the previous value downwards
airline_delays <- airline_delays %>% 
  mutate(airline = na_if(airline, '')) %>%
  fill(airline, .direction = c('down'))

```

```{r}
#Converting City rows into a column
airline_delays <- airline_delays %>% 
  gather(key = 'city', value = 'total.flights', 3:7)
#airline_delays

#separate by flight status and create column for total flights
flightstatus <- spread(airline_delays, status, total.flights) %>%
  mutate( total.flights = rowSums(across(where(is.numeric))))

#rename on time to not have space
flightstatus <- flightstatus %>%
  rename('on.time' = 4)

flightstatus

```

```{r}
#analyzing delay rates between the two airlines

#create columns for % of flights delayed and % of flights on-time

airline_delays_summarized <- flightstatus %>%
  group_by(airline, city) %>%
  #creating sums of the delayed and ontime flights to use for rate formulas
  summarise(total.delayed = sum(delayed), total.ontime = sum(on.time))%>%
  mutate(total = rowSums(across(where(is.numeric))),
       delay.rate = round(total.delayed / total*100, 2),
       ontime.rate = round (total.ontime / total*100, 2))

airline_delays_summarized


#creating Visuals
library(ggplot2)

#viewing delay rates by city per airline
ggplot(airline_delays_summarized, aes(fill= airline, y = delay.rate, x = city)) +
  geom_bar(position="dodge", stat = "identity")

#viewing ontime rates by city per airline
ggplot(airline_delays_summarized, aes(fill= airline, y = ontime.rate, x = city)) +
  geom_bar(position="dodge", stat = "identity")

```

## Conclusion

Based on the tidied data and the stacked bar charts, we can conclude that for every destination, AM West had a higher rate of delayed flights in comparison to Alaska.