---
title: "Project_2_Data_607"
author: "Natalie Kalukeerthie and Bishoy Sokkar"
date: "2024-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(1) Choose any three of the “wide” datasets identified in the Week 6 Discussion items. (You may
use your own dataset; please don’t use my Sample Post dataset, since that was used in your
Week 6 assignment!) For each of the three chosen datasets:

Create a .CSV file (or optionally, a MySQL database!) that includes all of the information
included in the dataset. You’re encouraged to use a “wide” structure similar to how the
information appears in the discussion item, so that you can practice tidying and
transformations as described below.

Read the information from your .CSV file into R, and use tidyr and dplyr as needed to
tidy and transform your data. [Most of your grade will be based on this step!]
Perform the analysis requested in the discussion item.

Your code should be in an R Markdown file, posted to rpubs.com, and should include
narrative descriptions of your data cleanup work, analysis, and conclusions.

(2) Please include in your homework submission, for each of the three chosen datasets:

The URL to the .Rmd file in your GitHub repository, and The URL for your rpubs.com web page.


### Dataset 1 - US Census: Population growth 2015 - 2020 of Southern States

This dataset contains population data for 4 states in the span of 6 years (2015 - 2020) and is sourced from the US Census Bureau.

We will analyse this data set in order to look for a things related to yearly population growth: 

1. Descriptive statistics of the data
2. Compare yearly growth between the states
3. See which year had the largest population growth

```{r}
#load the data
library(tidyverse)
library(openintro)
library(dplyr)
library(ggplot2)

population_data <- read.csv('https://raw.githubusercontent.com/nk014914/Data-607/main/Population_data.csv')

population_data

```

```{r}

#tidying the data

#Removing blank rows and renaming the columns to shorten for just the year
population_data <- population_data %>% 
  drop_na() %>%
  rename(state = Name, x2015 = X2015.Population, x2016 = X2016.Population, x2017 = X2017.Population, x2018 = X2018.Population, x2019 = X2019.Population, x2020 = X2020.Population)

#remove unused columns
population_data_subset <- subset(population_data, select= -c(FIPS,Abbreviations))

#transforming data - putting years into columns to make long data
population_data_subset <- population_data_subset %>% 
  gather(key = 'year', value = 'population', 2:7)

population_data_subset

#remove x's in year column
population_data_final <- population_data_subset %>%
  mutate(across(c('year'), substr, 2, nchar(year)))

population_data_final

```

### Analysis
```{r}
#First I need to make data as a numeric vector because my y axis was not in increasing order
transform(population_data_final, population = as.numeric(population))

#checking if column is now numeric
class(population_data_final$population)

```

```{r}
#Determine the summary statistics on population per year of the four states
population_summary <- population_data_final %>%
  group_by(year) %>%
  summarize(mean = mean(population), median = median(population), min = min(population), max = max(population), na.rm = T) %>%
  arrange(mean)

population_summary 

```

```{r}
#visualize growth of population per state
ggplot(data=population_data_final, aes(x=year, y= population, group = state)) +
  geom_line(aes(color=state))+
  geom_point(aes(color=state))
```

Viewing the yearly state populations plotted, we can see a few things:
- Georgia has the largest overall population every year, while South Carolina has the smallest (by a large difference in comparison to the other states.)

- Georgia and North Carolina are closer in yearly population than the other two states as their lines are very close to one another.

- Virginia seems to have the smallest yearly population growth as their line has the least amount of steepness in comparison to Virginia, North Carolina and Georgia.

```{r}
#Determining which year has the largest population growth

#sum of population per year
sum2015 = sum( population_data_final$population [ population_data_final$year==2015] )
sum2016 = sum( population_data_final$population [ population_data_final$year==2016] )
sum2017 = sum( population_data_final$population [ population_data_final$year==2017] )
sum2018 = sum( population_data_final$population [ population_data_final$year==2018] )
sum2019 = sum( population_data_final$population [ population_data_final$year==2019] )
sum2020 = sum( population_data_final$population [ population_data_final$year==2020] )

population_total = c(sum2015,sum2016,sum2017,sum2018,sum2019,sum2020)

#add column with years and create dataframe
year = c('2015','2016','2017','2018','2019','2020')
sum_population = as.data.frame(year)

sum_population$population_total <- population_total

sum_population

#calculate percent change per year over year, we will use the collapse package
library(collapse)

sum_population |> fmutate(growth = fgrowth(population_total))
```


## Conclusion

Based on the analysis, we can see the largest population growth of the four states combined is between 2018 and 2019 with a growth of around 1.29. The smallest growth was from 2015 to 2016 at 1.04. An interesting find I saw was that the growth percentage was steadily increasing each year until 2019-2020 where it dropped to 1.13 the second lowest population growth of all 6 years.The recent drop in population growth could be attributed to economical/financial issue pertaining to the year, or the impeding issues cause by the corona virus, however more research will be needed to confirm or deny these assumptions.


### Dataset 2 - Global Inflation Data

For this data set, we will look into the following:

1. Which region had the highest nad lowest CPI
2. Look at a single country (US) and any changes we may see per year along with its min/max

As suggested by Anna Moy and Mohammed Rahman, I used a csv called region in order to classify each country to a  region for simpler data categorization.

```{r}

#Load the csv
global <- read.csv('https://raw.githubusercontent.com/nk014914/Data-607/main/global_inflation_data.csv')
region <- read.csv('https://raw.githubusercontent.com/nk014914/Data-607/main/region.csv')

colnames(region) <- c("country_name", "region")
global_inflation <- merge(global, region, by = "country_name")

head(global_inflation)
```


```{r}
#Tidying the data

#drop unnecessary column and NAs
global_inflation <- subset(global_inflation, select = -c(indicator_name))
global_inflation <- global_inflation %>%
  select(country_name, region,X1980,X1980,X1981,X1982,X1983,X1984,X1985,X1986,X1987,X1988,X1989,X1990,X1991,X1992,X1993,X1994,X1995,X1996,X1997,X1998,X1999,X2000,X2001,X2002,X2003,X2004,X2005,X2006,X2007,X2008,X2009,X2010,X2011,X2012,X2013,X2014,X2015,X2016,X2017,X2018,X2019,X2020,X2021,X2022,X2023,X2024) %>%
  pivot_longer(`X1980`:`X2024`,names_to = "year", values_to = "CPI_inflation", values_drop_na = TRUE) 


#remove x's in year column
global_inflation_final <- global_inflation %>%
  mutate(across(c('year'), substr, 2, nchar(year)))

global_inflation_final

```

### Analysis

```{r}
#Determine min and max of the regions, sorting by desc
global_max <- global_inflation_final %>%
  group_by(region) %>%
  summarize(max_cpi = max(CPI_inflation)) %>%
  arrange(desc(max_cpi))

global_min <- global_inflation_final %>%
  group_by(region) %>%
  summarize(min_cpi = min(CPI_inflation)) %>%
  arrange(min_cpi)

global_min
global_max
```

```{r}
#Filter out the US to look at its min/max and determine its lowes CPI avg by sorting

#Highest/lowest CPI year
us_CPI <- global_inflation_final %>%
  filter(country_name == "United States") %>%
  mutate(min = min(CPI_inflation), max = max(CPI_inflation)) %>%
  arrange(desc(CPI_inflation))
us_CPI

#CPI by year
us_CPI_year <- global_inflation_final %>%
  filter(country_name == "United States") %>%
  mutate(min = min(CPI_inflation), max = max(CPI_inflation)) %>%
  arrange(desc(year))

us_CPI_year
```

## Conclusion

After tidying and analyzing the data, the highest CPI for all regions is 65374 in the South/Latin America region. The lowest CPI is seen in Africa at -72.7. The drastic difference between the CPIs can be attributed to outliers. When looking at just the US in CPI descending order, we see 2009 had the lowest CPI (min) at -0.3, while 1980 has the highest CPI (max) at 13.5. Looking at year over year and outside 1981 and 1980, the US CPI did not fluctuate much and remained relatively constant.


### Dataset #3 - CPI data for January 2024

Dataset: ["Bureau of Labor Statistics - CPI Report for Jan2024"](https://www.bls.gov/cpi/tables/supplemental-files/home.htm)\


```{r}
# Load the data
library(tidyverse)

CPI_Jan <- read.csv("https://raw.githubusercontent.com/sokkarbishoy/DATA607/main/news-release-table1-202401.csv")

head(CPI_Jan)

```


```{r}
#We start tidying the data removing the first two rows, last three rows and empty rows

CPI_Jan <- CPI_Jan[-c(1:5, 16, 25, 45:49), ]

tail(CPI_Jan)

```

```{r}
#change column names

new_columns <- c("Indent_Level",

                 "Expenditure_category",

                 "Relative_importance_Dec_2023",

                 "Jan_2023_Unadjusted_indexes", 

                 "Dec_2023_Unadjusted_indexes", 

                 "Jan_2024_Unadjusted_indexes",

                 "Jan_2023_Jan_2024_Unadjusted_percent_change",

                 "Dec_2023_Jan_2024_Unadjusted_percent_change", 

                 "Oct_2023_Nov_2023_Seasonally_adjusted_percent_change",

                 "Nov_2023_Dec_2023_Seasonally_adjusted_percent_change", 

                 "Dec_2023_Jan_2024_Seasonally_adjusted_percent_change")

colnames(CPI_Jan) <- new_columns

 

#change the type into INT

CPI_Jan$Indent_Level <- as.integer(CPI_Jan$Indent_Level)

head(CPI_Jan)

```

Finally we change the type of each columns to INT and Numeric

```{r}

CPI_Jan <- transform(CPI_Jan, 

          Indent_Level = as.integer(Indent_Level),

          Relative_importance_Dec_2023 = as.numeric(Relative_importance_Dec_2023),

          Jan_2023_Unadjusted_indexes = as.numeric(Jan_2023_Unadjusted_indexes),

          Dec_2023_Unadjusted_indexes = as.numeric(Dec_2023_Unadjusted_indexes),

          Jan_2024_Unadjusted_indexes = as.numeric(Jan_2024_Unadjusted_indexes),

          Jan_2023_Jan_2024_Unadjusted_percent_change = as.numeric(Jan_2023_Jan_2024_Unadjusted_percent_change),

          Dec_2023_Jan_2024_Unadjusted_percent_change = as.numeric(Dec_2023_Jan_2024_Unadjusted_percent_change),

          Oct_2023_Nov_2023_Seasonally_adjusted_percent_change = as.numeric(Oct_2023_Nov_2023_Seasonally_adjusted_percent_change),

          Nov_2023_Dec_2023_Seasonally_adjusted_percent_change = as.numeric(Nov_2023_Dec_2023_Seasonally_adjusted_percent_change),

          Dec_2023_Jan_2024_Seasonally_adjusted_percent_change = as.numeric(Dec_2023_Jan_2024_Seasonally_adjusted_percent_change))

```

Analyzing the data:

in the next lines, we will answer the question: which categories saw the biggest from December to Jan. Amother part of tidying the data will be using the filter function to analyse the subgroups only which will give us accurate data.

Data shows that commodities expenditures such as Cereals, bakery products, as well as Energy and gasoline saw the least change.\

```{r}

CPI_Jan %>%

  filter(Indent_Level == 3) %>%

  arrange(Dec_2023_Jan_2024_Seasonally_adjusted_percent_change)

```

To view expenditures that saw the most change we can use the following code.

```{r}

CPI_Jan %>%

  filter(Indent_Level == 3) %>%

  arrange(desc(Dec_2023_Jan_2024_Seasonally_adjusted_percent_change))

```

Below we analyze the major categories (indent level = 3);

```{r}

CPI_Jan %>% 

  filter(Indent_Level == 3) %>%

  arrange(desc(Relative_importance_Dec_2023))

```

## Conclusion
Conclusion: CPI data contains many more insights that I can be analyzed. Here are the questions we answered in the analysis above:

Which category saw the least change from Dec 2023 - Jan 2024? Ans: Fuel oil, Used cars and trucks, and Motor fuel.
A:  Fuel oil, Used cars and trucks, and Motor fuel.

Which category saw the biggest change from Dec 2023 - Jan 2024? Ans: Utility Gas services, electricity, and transportation services.
A:Utility Gas services, electricity, and transportation services.

Which category saw the biggest change from Jan 2023 - Jan 2024? Ans: Transportation, Tobacco and smoking products, and Shelter. 
A: Transportation, Tobacco and smoking products, and Shelter. 