---
title: "project-1_Data_607"
author: "Natalie Kalukeerthie"
date: "2024-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605


### Loading the data

```{r}
#load the data
library(tidyverse)
library(openintro)
library(readr, quietly = TRUE)
library(stringr, quietly = TRUE)

chess_raw <- 'https://raw.githubusercontent.com/nk014914/Data-607/main/project_1_tournamentinfo.txt'

#place in dataframe
df <- read_lines(chess_raw)

head(df)
```


### Configuring the Data

```{r}
#locate the pattern occurences
x0 = 0
x1 = unname(str_locate_all(pattern = '\\|', df[5])[[1]][1,1])
x2 = unname(str_locate_all(pattern = '\\|', df[5])[[1]][2,1])
x3 = unname(str_locate_all(pattern = '\\|', df[5])[[1]][3,1])
x4 = max(nchar(df))

#sequence and group
seq1 = seq(5, 196, 3)
seq2 = seq(6, 196, 3)
group1 = df[seq1]
group2 = df[seq2]


```


### Grouping the Data
```{r}

#grouping for each column

#player names
name = substr(group1, x1 + 1, x2 - 2)
PlayerName = str_trim(name)

#player state
state = substr(group2, x0, x1 - 1)
State = str_trim(state)

#total points
totalpts = substr(group1, x2 + 1, x3 - 1)

#gpre-ratings
pre = substr(group2, x1 + 1, x2 - 1)
pre = str_extract(pre, ': *\\d{2,}')

#add into df
chess_scores = data.frame(PlayerName, State)
chess_scores$TotalPts = sprintf("%.1f", as.numeric(totalpts))
chess_scores$PreRating = as.integer(str_extract(pre, '\\d{2,}'))
```


### Calculating Opponent's Avg Pre Rating

```{r}
opp = substr(group1, x3 + 1, x4)
opp = str_extract_all(opp, '\\b\\d{1,}')
opp = as.matrix(opp)

avgoppprerate = function(y, z){
  x = y[z]
  
  for (a in x) {
    rate = 0
    c = 0
    for (b in a) {
      c = c + 1
      rate = rate + chess_scores$PreRating[as.numeric(b)]
    }
    rate = round(rate/c)
  }
  return(rate)
}

chess_scores$AvgOppPreRating = apply(opp, 1, avgoppprerate)
```


### Converting txt to CSV

```{r}
head(chess_scores)

#convert txt into csv

write.csv(chess_scores, "chesstournamentinfo.csv")

```